### 2010-04-25 product-graphics.tex

\documentclass[letterpaper,oneside]{scrartcl}
\usepackage{fullpage}
\usepackage[utf8]{inputenc}
\usepackage[pdftex]{graphicx}
\usepackage{tikz}
\DeclareGraphicsExtensions{.png,.pdf}
\graphicspath{{plots/}}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{url}
\usepackage[round,sectionbib]{natbib}
\bibliographystyle{abbrvnat}
\usepackage[small]{caption}
\usepackage[small]{titlesec}
\renewcommand\familydefault{bch}

\definecolor{female}{HTML}{EF8A62}
\definecolor{male}{HTML}{67A9CF}
\definecolor{not-too-happy}{HTML}{FC8D59}
\definecolor{pretty-happy}{HTML}{FFFFBF}
\definecolor{very-happy}{HTML}{91CF60}
\definecolor{grey30}{HTML}{4D4D4D}
\newcommand{\key}[1]
  {\protect \tikz{\fill[#1] rectangle (1ex,1ex);}}

\title{Product graphics}
\author{Hadley Wickham and Heike Hofmann}
\begin{document}
\maketitle

\section{Introduction}

% JASA methods
% Infovis?

% Aims of this paper:
%
%  * emphasise the connection between area graphics and factorising
%    probabilities
%
%  * pull all area plots on a common footing - provide a grammar for
%    categorical graphics

Area plots, like bar charts and mosaic plots, are the graphical equivalent of contingency tables, displaying how a total is broken down.

Display of discrete distribution, pmf, contingency table.

This paper presents a unifying framework for area plots, showing how many existing plots are interrelated, and by drawing an

The paper proceeds as follows. Section~\ref{sec:primitives} defines the simple one- and two-d primitives used for all more complicated graphics. Section~\ref{sec:combination} provides a mathematical framework that shows how these simple primitives can be combined to display higher-d tables. Section~\ref{sec:variations} shows how these plots can be adapted to deal with missing values, weights and continuous data.

To illustrate these ideas, the same data for all examples. It's a small sample
of variables related to happiness from the general social survey ({\sc gss}).
The {\sc gss} is a yearly cross-sectional survey of Americans. Combining
surveys from 1972 to 2006 gives 51\,020 observations. Of the over 5\,000
variables, we select eight related to happiness, as described in
Table~\ref{tbl:happy}. The data is included in the supplemental materials.

\begin{table}
  \begin{center}
  \begin{tabular}{llp{8cm}}
    \toprule
    Variable name & Description & Values \\
    \midrule
    {\sf age} & age in years & 18--89 \\
    {\sf degree} & highest education & lt high school, high school, junior college, bachelor, graduate \\
    {\sf finrela} & relative financial status & far above average, above average, average, below average, far below average \\
    {\sf happy} & happiness & \key{very-happy} very happy, \key{pretty-happy} pretty happy, \key{not-too-happy} not too happy \\
    {\sf health} & health & excellent, good, fair, poor \\
    {\sf sex} & sex & \key{female} female, \key{male} male\\
    {\sf wtsall} & probability weight & 0.43--6.42 \\
    {\sf year} & year of survey & 1972--2006 \\
    \bottomrule

  \end{tabular}
  \end{center}
  \caption{Description of sample data.  Common plot colours are shown next to respective levels.}
  \label{tbl:happy}
\end{table}

\section{Graphical primitives}
\label{sec:primitives}

Tilings of the plane. In general, we could also think of partitions of higher-dimensional spaces (e.g. 3d or 4d), but we will need to project these down to two for viewing anyway, so there is little disadvantage to working directly in 2d. To derive the graphical primitives we start with three constraints:

\begin{itemize}
  \item Area must be {\bf proportional} to value. Because the total area for a graphic is usually constrained, an area plot serves to partition up this total, displaying proportions rather than counts.

  \item Partitions must be {\bf rectangular}. It is easier to compare the areas of simple shape (i.e. rectangle vs polygon), and even better when one the length of one border is fixed, so we only need compare the length of the other \citep{cleveland:1984}. Rectangular partitions are also recursive in the sense that we can always tile a rectangle with smaller rectangles. This is not true for most shapes.

  \item Partitions must be {\bf disjoint}. To be able to see the complete area of each rectangle they must be non-overlapping, and each split must be nested within the shapes of the previous level. Note that we do not impose the constraint that the tiling be space-filling. Space filling methods use will ensure that areas are larger, but make it harder to compare magnitudes because scales no longer align.

\end{itemize}

\noindent Each of these three constraints can be profitably relaxed, as detailed in Section~\ref{sec:variations}.

The following sections describe the partitions generated by these constraints. Section~\ref{sub:part-1d} discusses 1d partitions, bars, spines and tiles, and  Section~\ref{sub:part-2d} discusses the 2d partition, the fluct.

% Let $a(i)$ give the area assigned to category $i$, which has proportion $f(i)$. Since we are using rectangles, $a(i) = k \dot a(i) = k \dot w(i) \times h(i)$ where $w$ and $h$ represent width and height respectively. Certain partitions also satisfy the constraint that $h(i) = \sqrt{k} f(i)^x$ and $w(i) = \sqrt{k} f(i)^y$ where $x + y = 1$. This is useful because it is only necessary to inspect one border to compare magnitudes.

\subsection{1d primitives}
\label{sub:part-1d}

There are three 1d primitives, as shown in Figure~\ref{fig:part-1d}:

\begin{itemize}

  \item {\bf bar}s: height is proportional to value, width equally divides space. Bars allow comparison between absolute numbers, but are not space filling. Bars are not space filling, occupying $\sum(p_i / \max(p_i))$ of the total area. Bars can be arranged horizontally (``hbar'') or vertically (``vbar'').

  \item {\bf spine}s: width is proportional to value, height is fixed. Splines are space filling. Splines can be arranged horizontally (``hspine''), vertically (``vspine''), or can automatically pick their orientation (``spine'').

  \item {\bf tile}s: tile the plane with rectangles, trying to keep the aspect ratio of each rectangle close to 1. There are no restrictions on height or width, just the overall area. This is partitioning used by the squarified treemap \citep{bruls:1999}.

\end{itemize}

Each of these three displays has different strengths and weaknesses.  The perceptual task associated with bar charts is the simplest, but they do not occupy all the area. Spines and tiles are harder to read, but occupy the complete space and work better recursively \citep{heer:2010}.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.33\linewidth]{part-hbar}%
    \includegraphics[width=0.33\linewidth]{part-hspine}%
    \includegraphics[width=0.33\linewidth]{part-treemap}

  \caption{1d partitions showing the distribution of happiness.  From left to right: bars, spines and tiles.}
  \label{fig:part-1d}
\end{figure}

The appearance of bars and splines is identical for uniformly distributed data.

\subsection{2d primitives}
\label{sub:part-2d}

We are currently aware of one 2d primitive, the {\bf fluct}, which has both height and width proportional to the square root of the proportion. Flucts allow comparisons of relative magnitudes in two directions. Each rectangle is arranged on a regular grid formed by the levels of the two variables, and so is not space filling.

One special case of the fluct is the equal bin size plot which occurs when the two variables are jointly uniformly distributed, usually as a result of conditioning. The equal bin size plot is useful to visualise missing combinations, or when supplemented with a third variable, to visualise the distribution conditional on two other variables.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.5\linewidth]{part-fluct}%
    \includegraphics[width=0.5\linewidth]{part-fluct-cond}
  \caption{(Left) Distribution of age (in decades) and marital status. (Right) Equal bin-size plot showing health status and survey year. Health status was not recorded for three years.}
  \label{fig:fluct}
\end{figure}

\section{Factorising area plots}
\label{sec:combination}

To construct plots for discrete pdfs with more than two dimensions, we need a mechanism to factorise them into low dimensional products. This factorisation has three roles:

\begin{itemize}
  \item as products of marginal and conditional distributions
  \item as areas in plots
  \item as sums of parameters in log-linear models
\end{itemize}

Let $f(x, y, z)$ be the 3d pmf function. We can can write this pmf as product of marginals and conditionals in many different ways.  We can decompose product graphics in a similar way.

\begin{itemize}
  \item $f(x, y, z) = f(x, y | z) f(z)$
  \item $f(x, y, z) = f(x | y, z) f(y, z)$
  \item $f(x, y, z) = f(x | y, z) f(y | z) f(z)$
\end{itemize}

% It can be useful to display partial decompositions, i.e. just $f(x, y | z)$, which is still a 3 variable pmf, and is defined as
%
% \[ f(x, y | z) = \frac{f(x, y, z)}{f(z)} \]
%
% so that each category of z sums to one, and for any fixed value of z, $f(x, y | z)$ is a valid pmf.
%
% \[ \forall b \sum_a f(a | b) = 1 \]
%
% Because of their hierarchical nature, an area plot of $f(x, y, z)$ also shows $f(y, z)$ and $f(z)$.

Based on the modelling specification of \citet{wilkinson:1973} we define an algebra for describing these tables.

\begin{itemize}
  \item \verb|~ a| represents $f(a)$,
  \item \verb|~ a + b| represents $f(a, b)$
  \item \verb!~ a | b! represents $f(a | b)$
  \item \verb!~ . | a! is a special case of $f(a) = \frac{1}{\#A}$.
\end{itemize}

By convention we will write the top most level of the product plot on the left, and lower levels to the right.  While $f(x, y, z)$ and $f(z, y, x)$ display the same counts, their interpretation is different because of which marginal tables you see.

Figure~\ref{fig:fact-simple} uses the distribution of sex and happiness to illustrates these ideas. From left to right, we have plots of \verb|~ happy|, \verb!~ happy | sex! and their product \verb|~ happy + sex|.  Observe that the heights and widths of the rectangles are multiplied in the same way as the components of the pmf.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.3\linewidth]{fact-happy} $\times$ %
  \includegraphics[width=0.3\linewidth]{fact-happy|sex} = %
  \includegraphics[width=0.3\linewidth]{fact-happy-sex}

  \includegraphics[width=0.3\linewidth]{fact-happy-2} $\times$ %
  \includegraphics[width=0.3\linewidth]{fact-happy|sex-2} = %
  \includegraphics[width=0.3\linewidth]{fact-happy-sex-2}

  \caption{Plots of the distribution of happiness and sex (\key{male} male, \key{female} female).}
  \label{fig:fact-simple}
\end{figure}

% To visualise higher-dimensional pmfs, we recursively partition using the previously described primitives. That is, to display $f(x, y, z)$ we first create a partition based on $f(z)$ then partition each of those pieces based on $f(y | z)$ and again by $f(x | y, z)$.

Figure \ref{fig:recursive} illustrates this process with a graphic that explores whether, conditional on marital status, there is a difference in happiness between men and women.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.33\linewidth]{part-comb-1}%
    \includegraphics[width=0.33\linewidth]{part-comb-2}%
    \includegraphics[width=0.33\linewidth]{part-comb-3}
  \caption{Conditional on marital status, are men or women happier?  This figure shows the construction of $\sim happy + sex + marital$ from (left to right) vspines by marital status, vspines by marital status and sex, and hspines by marital status, sex and happiness.  For all levels of marital status, men are slightly less happy.}
  \label{fig:recursive}
\end{figure}

Different combinations of partitions reveal different features of the data. Take for example, the distribution of age and marital status, as shown in Figure~\ref{fig:fluct}. Instead of visualisation the joint distribution with a fluct, we could focus on the conditional distribution of marital status, given age, or vice versa. Figure~\ref{fig:marital} shows two ways to do this. The left plot shows $\sim marital + age$ with a vspine nested in a hspine, and the right plot $\sim age + martial$ with a hbar nested in a vspine. These displays show the same data but are arranged in such a way to make different comparisons easier.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.5\linewidth]{part-marital-1}%
    \includegraphics[width=0.5\linewidth]{part-marital-2}
  \caption{(Left) The joint distribution of marital status and age (in decades) partitioned by a vspine and hspine.  (Right) The joint distribution of age and marital status partitioned by a vspine and hbar.}
  \label{fig:marital}
\end{figure}

Conditioning is also an important tool, because it allows one to remove relationships that are well-known or uninteresting. Figure~\ref{fig:part-cond} uses a fluct and a vspine to explore the relationship between happiness, health and financial status. The left plot displays raw proportions, showing that most people are in good health and average financial standing. However, it is difficult to see how happiness varies within these conditions because the perceptual task is comparing areas. Conditioning on financial status and health produces the plot on the right (an equal bin size plot) and makes it easier to see the conditional distribution of happiness, because the perceptual task is simpler: comparing length on a common non-aligned scales.  Here we can see people tend to be happier the healthier and richer they are.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.5\linewidth]{part-fluctuation}%
    \includegraphics[width=0.5\linewidth]{part-equal-area}
  \caption{Two displays of $\sim happy + health + finrela$.  (Left) Unconditional values partitioned with a vspine and fluct.  (Right) Same partitions, but health and relative financial status conditioned to uniform distribution.  We can no longer see the joint distribution of health and financial status, but it is much easier to see the conditional distribution of happiness: healthier and richer people are happier. (Maybe money does buy happiness?).  (\key{not-too-happy} Not too happy, \key{pretty-happy} pretty happy, \key{very-happy} very happy)}
  \label{fig:part-cond}
\end{figure}

% http://peltiertech.com/Utility/ClusterStackUtility.html

\subsection{Existing plot types}

% Other papers to cite:
%
% Generalised treemaps: \citep{vliegen:2006}
% Cocktailmaps: \citep{ahlberg:1996}
% Extension to mosaic: \citep{friendly:1999,hofmann:2000}
% Connection to models: \citep{theus:1999,hofmann:2001}

The sequence of partitioning choices defines different named area plots:

\begin{itemize}
  \item {\bf Barchart} (1d). 1 hbar.
  \item {\bf Column chart} (1d). 1 vbar.
  \item {\bf Spineplot} (1d). 1 vspine.

  \item {\bf Stacked} barchart (2d). 1 hbar and 1 vspine.
  \item {\bf Nested} barchart (2d).  2 hbars. \citep{peltier:2009}
  \item {\bf Fluctuation} diagram (2d): fluct.

  \item {\bf Equal bin size} plot (3d): fluct and vspine, provided the first two variables are uniform.

  \item {\bf Mosaic} plot (nd).  Alternating hspines and vspines.  \citep{hartigan:1984,hartigan:1981,friendly:1994,hofmann:2003}
  \item {\bf Double-decker} plot (nd).  $n-1$ hspines and 1 vspine. \citep{hofmann:2001}
  \item {\bf Treemap} (nd): n spines. \citep{shneiderman:1992}
  \item {\bf Squarified treemap} (nd): n tiles. \citep{bruls:1999}

\end{itemize}

Trellis graphics \citep{becker:1996}, as known as lattice or facetted or
conditioned graphics are another related display. They uses categorical
variables to generate multiple panels, each containing a plot of the subset of
the data. Trellised area plots also fall into our framework and can be
created by conditioning on the trellising variables.

These techniques are also closely related to dimensional stacking \citep{leblanc:1990}

\section{Modelling}
\label{sec:models}

\section{Variations and extensions}
\label{sec:variations}

On top of this basic framework, it is useful to consider a few variations and extensions.  Three are based on variations of the underlying data (missing data, weighted data, and continuous data), and three are based on relaxing the display constraints (area not proportional to weight, non-disjoint partitions, non-rectangular partitions).

\subsection{Missing values}
\label{sub:missing_values}

\subsection{Weighting}
\label{sub:weighting}

We have assumed that the probabilities represent counts, but without loss of generality, we can use any set of non-negative, additive weights. For example, in the happy dataset, the {\sf wtssall} variable gives analytic survey weights. Figure~\ref{fig:weighting} shows the difference between the weighted and unweighted distributions of age and sex. The distribution is subtly different. \citep{unwin:2007} describes the applications of weighted plots in more detail.

To express that we are displaying weights, rather than counts, we extend the formula syntax by putting the weighting variable on the left-hand side: $wtssall \sim sex + age$ vs. $ \sim sex + age$.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.5\linewidth]{wt-count}%
    \includegraphics[width=0.5\linewidth]{wt-wtssall}%
  \caption{Joint distribution of age and sex (\key{male} male, \key{female} female). (Left) counts and (right) probability weights.}
  \label{fig:weighting}
\end{figure}

\subsection{Continuous data}
\label{sub:continuous_data}

These techniques can also be used with continuous data, if the data is binned to create a categorical variable. There are many different possible choices of binning algorithms, but we will focus on two:

\begin{itemize}
  \item Intervals of equal width.
  \item Intervals containing equal numbers of points.
\end{itemize}

This gives rise to the histogram and spinogram, continuous analogues of the bar chart and spine chart. A long standing tradition is that no gaps are displayed between adjacent rectangles when used for continuous data. Two examples are shown in Figure \ref{fig:cont-examples}.

\begin{figure}[htbp]
  \centering
  \caption{Area plots of originally continuous data}
  \label{fig:cont-examples}
\end{figure}

Example exploring joint and conditional distributions of age and year.

% An alternative approach is use density estimates (or conditional density estimates) and plot with continuous area plots.  As in: Hofmann, H., Theus, M. (2005), Interactive graphics for visualizing conditional distributions, Unpublished Manuscript.

\subsection{Area not proportional to weight}

It can be useful to violate the constraint that area should be proportional to weight to distinguish between zeros and very small values. A zero weight should have zero area, but giving it positive area can be useful, so you can actually see it! In general, it's useful to constrain all areas to be above a certain minimal perceptible size.  Areas which are constrained in such a way should have a visual flag (such as a difference colour) so ensure that the reader knows that the relationship between area and value has been violated.

On the other end of the spectrum, it can be useful to constrain put a cap on the largest possible values \citetext{Antony Unwin, priv. comm.}. By controlling this cap it is possible to zoom in on the smallest values. - called censored zooming by Antony and Martin.

Other non-linear transformations may also be useful. For example, we could display the square root of the weights, to stabilise the variance of the areas. Tukey applied this technique to histograms to create rootograms.

If values span multiple orders of magnitude it may be useful to map log value to area.

\subsection{Non-disjoint partitions}

The cascaded treemaps of \citet{lu:2008} is an idea that illustrates how the violation of containment can be productive.  In the cascaded treemap, each level is slightly offset from the one above to create a pseudo-3d perspective.
This makes it easier to see all the levels of the hierarchy, not just the lowest level.  Figure~\ref{fig:cascade} shows an example of how cascading can help illuminate the structure of a complex mosaic plot.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.5\linewidth]{sex-marital-happy}%
    \includegraphics[width=0.5\linewidth]{cascade}
  \caption{A mosaic plot of happiness by martial status and sex. (Left) Coloured by happiness (\key{not-too-happy} Not too happy, \key{pretty-happy} pretty happy, \key{very-happy} very happy).  (Right) A cascaded view helps show how the plot is built up.}
  \label{fig:cascade}
\end{figure}

\subsection{Non-rectangular partitions}

A pie chart is a popular method of displaying proportions, but is not a rectangular partition. However, it turns out that there is a simple relationship between product plots and pie charts: an pie chart is an hspine drawn in polar coordinates with the x coordinate mapped to angle and the y coordinate to radius. Many other displays that use circular arcs turn out to be special cases of product plots, drawn in polar coordinates:

\begin{itemize}
  \item Fourfold displays \citep{friendly:1995}, wind rose, sector graphic: hbar.

  \item Bullseye chart: hspine.

  \item Doughnut plot: vspine + hspine.

  \item Circular bar chart or racetrack plot: vbar.

  \item Infoslices \citep{andrews:1998}, nested vbar. But they only use half of the disk, and are specialised for highly nested data.

  \item Nightingales's coxcomb, is very similar to vspine / hbar in polar coordinates, except that the pie slices are overlapping and so violates the constraint of disjoint area.
\end{itemize}

\noindent Eight variants are displayed in Figure~\ref{fig:polar}.

To make this work more generally, the y-variable (mapped to radius) should be square-root transformed to ensure that that counts stay proportional to areas. The radial displays of \citet{stasko:2000} and fanlens of \citet{lou:2007} deliberately choose not to do this to highlight the outer levels.

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.25\linewidth]{hb-vb-cartesian}%
    \includegraphics[width=0.25\linewidth]{hb-vs-cartesian}%
    \includegraphics[width=0.25\linewidth]{hs-hs-cartesian}%
    \includegraphics[width=0.25\linewidth]{hs-vs-cartesian}

    \includegraphics[width=0.25\linewidth]{hb-vb-polar}%
    \includegraphics[width=0.25\linewidth]{hb-vs-polar}%
    \includegraphics[width=0.25\linewidth]{hs-hs-polar}%
    \includegraphics[width=0.25\linewidth]{hs-vs-polar}

    \includegraphics[width=0.25\linewidth]{hb-vb-polar-2}%
    \includegraphics[width=0.25\linewidth]{hb-vs-polar-2}%
    \includegraphics[width=0.25\linewidth]{hs-hs-polar-2}%
    \includegraphics[width=0.25\linewidth]{hs-vs-polar-2}

  \caption{(Top) Area graphics in Cartesian coordinates. (Mid) Area graphics in polar coordinates. From left to right: hbar + hbar, vspline + hbar, hspine + hspine, hspine + vspine. (Bottom) Product graphics rotated 90 degrees before being converted to polar coordinates: vbar + vbar, hspline + vbar, vspine + vspine, vspine + hspine. (\key{male} male, \key{female} female)}

  \label{fig:polar}
\end{figure}

A number of non-rectangular treemaps have been proposed, such as the circular treemaps of \citet{wetzel:2008}, the space-filling curves of \citet{wattenberg:2005} and the voronoi treemaps of \citet{balzer:2005}.  However, because it is so much more difficult to compare the areas of arbitrary polygons, these approaches tend to be attractive rather than useful.

\section{Conclusions}

% bibtool -x product-graphics -o references.bib
\bibliography{references}
\end{document}