### 2011-06-20 reviews/response.txt
Add reviews

# Summary review

The reviewers judge Product Plots to be a useful and parsimonious model
for constructing a large variety of plots and potentially also an example
that might spur further research in formal models of visualization
construction. All reviewers agree that this work makes a substantive
contribution to InfoVis research and argue for acceptance.

That said, reviewers also raise a number of issues that the authors
should address prior to publication. Most notable among these is the
treatment of related work. R1 and R4 in particular list elements of prior
work that should be considered and discussed. I recommend that the
authors introduce a dedicated related work section in their next
revision.

In addition, the reviewers make a number of helpful suggestions that
might inform future drafts. I strongly encourage the authors to revise
the paper accordingly.

# Major changes

* Related work section & missing references
* Improve justification of rectangles using references provided.

# Review 2

* First, related work is distributed throughout the paper. A consolidated
related work section addressing similar work and clearly delineating the
contributions of the present work is needed. Other visualization models
of interest include Slingsby et al's HiVE, Wilkinson's Grammar of
Graphics, Mackinlay's APT, and Stolte et al's Polaris. A detailed
comparison with such prior work is a pre-requisite to publication. How do
such models compare in terms of expressiveness and notational efficiency?

* I'm curious how much effort is necessary to use product plots in
practice. Some examples of statements in R and the corresponding plots
should suffice (and potentially examples of an interactive UI for product
plot construction).

* Is anyone using the system to perform analyses? While not strictly
necessary, example analysis case studies supported by product plots might
help convey the value of the model.

* "the areas of rectangles are easier to compare than areas of other shapes
[5-7]" -> This is somewhat glibly stated. A quick review of the cited
papers does not support the claim: [5] does not assess area empirically;
[6] assesses circular and arbitrary polygonal areas, but not rectangles;
[7] compares circles and rectangles, but the difference does not appear
to be statistically significant. Later in the paper, there is a similar
critique of non-rectangular polygons using the results of [6]. I don't
think this is appropriate -- [6] uses irregular polygons, whereas
techniques like voronoi treemaps generate convex polygons with average
bounding box aspect ratios near unity. Thus the argument here is also
unsubstantiated. It may well be that rectangles are better, but this has
not been convincingly demonstrated. I recommend the authors consider
dropping these perceptual arguments.

# Review 4

* A surprising omission is the work of Wilkinson (1999, 2005) on The
 Grammar of Graphics. Wilkinson introduced the idea of combining graphical
 elements through products of sets (called facets) and embedding these in
 different coordinate systems. In fact, Figure 12 echoes several of the
 figures in the Coordinates chapter of that book. Wilkinson was also the
 first to draw a link between the recursive partitioning in treemaps and
 mosaics and a set product operator. As he explains in Chapter 16 of the
 second edition, the algebra for specifying visualizations of product
 models maps one-to-one to the most common factorial experimental design
 algebra for log-linear models or ANOVA. In any case, the framework in
 this paper appears closer to that of Wilkinson's than to any of the other
 papers cited.

 Bottom of page:
 Using the word "all" is a bit strong here. Better to delete the word and
 confine the discussion to "area" plots. There are exceptions, as the
 authors point out later, so avoid the inclusive term.

 The claim for rectangles at the bottom of page 1 is not supported by
 research. The authors, like many statisticians and visualization
 researchers, tend to cite statisticians and computer scientists for
 supportive psychological principles; however, the perceptual and
 cognitive research by non-psychologists tends to have serious
 methodological flaws. I've provided some relevant citations below. There
 are significant biases in area judgments that are exacerbated by extreme
 aspect ratios (something even Kong, Heer & Agrawala noticed in Treemaps).
 The perceptual combining rule for rectangles appears to be more Attneave
 (absolute metric) than product (multiplicative), although see Algom et
 al. for support for the multiplicative rule. The authors need to rephrase
 this statement, eliminating the word "must."

 Also, Simken and Hastie, in one of the few perception articles published
 by psychologists in a statistical journal, demonstrated conclusively that
 pie charts outperform rectangular area charts for proportion of whole
 data. There are some exceptions to their findings, but statisticians need
 to understand that they are largely wrong in their scorn for pie charts.

 We can tile triangles with triangles. And these are "simpler" than
 rectangles. One can also tile the sphere with triangles, which cannot be
 done with rectangles. You need to rethink this assertion.

 References:
 Algom, D., Wolf, Y., & Bergman, B. (1985). Integration of stimulus
 dimensions in perception and memory: Composition rules and psychophysical
 relations. Journal of Experimental Psychology: General, 114, 451-471.

 Anderson, N.H. and Weiss, D.J. (1971). Test of a multiplying model for
 estimated are of rectangles. American Journal of Psychology, 84, 4,
 543-548.

 Borg, I. and Leutner, D. (1983). Dimensional models for the perception of
 rectangles. Perception & Psychophysics, 34, 257-267.

 Kidman, G. (1997). Area integration rules. In F. Biddulph & K. Carr
 (Eds.), People in mathematics (Proceedings of the 20th annual conference
 of the Mathematics Education Research Group of Australasia (Vol. 1, pp.
 271 - 277).

 Mates, J., Di Maio, V., and Lansky, P. (1992). A model of the perception
 of area. Spatial Vision, 6, 101-116.

 Morgan M.J. (2005). The visual computation of 2-D area by human
 observers. Vision Research, 45(19), 2564-70.

# Review 1
 Finally, I wonder if the authors considered any other graphical
 composition method than nesting areas? While nesting areas allowed them
 to encompass a number of existing plot types, other composition methods
 might be more useful for extension to higher dimensions. For example,
 what about simply showing a separate plot for each component?

 -label all appropriate plots in the paper with the corresponding
 distribution decomposition (directly label, not just mention in the text
 or in the caption)
 -flesh out section 3.1. Assuming that infovis readers will not know what
 conditional and marginal distributions are, the authors should
 demonstrate how they are computed from the joint distribution. Providing
 the graphical analogs to the tables in Figure 4 would be extremely
 helpful. (I realize that a marginal and a conditional distribution from
 Figure 4 are represented in Figure 5, but the connection is never made
 explicitly in the text. Some hand holding would be very helpful here.)
 Hand hold better the analytic questions that each distribution helps
 answer. E.g. "What proportion of males are very happy? The bottom left
 entries tells us it's 31%."

 Minor comments----

 Why squarified treemaps? Heer work showing non-square treemaps are easier
 to read.

 Section 3.1

 Its not clear from the wording in the first paragraph that the marginal
 and conditional distributions contain less information than the joint
 distribution. If you're going to assume that your readers don't know what
 the marginal and conditional distributions are (plausible for the InfoVis
 crowd), then it would make sense to briefly describe how they are
 computed from the joint distribution.

 Fix the formatting of the tables in Figure 4.

 Clarify in the last paragraph that in addition to the 3 factorings,
 permutations of the variables leads to more variations.

 "It remains to find an [sic] graphical analog of this multiplication."
 seems unnecessarily dramatic. We already know you found one.

 Section 3.2

 Label Figure 5 with the distribution forms, f(happy), f(happy|sex),
 f(happy, sex), rather than just mention them in the text.

 Figure 7 shows, f(marital | age) f(age) and f(age | marital) f(marital),
 right? Say this in the text and label figures.

 "the relationship between proportion and area is only constant within a
 level" -- is there a better way to say this? I think you mean level as in
 nesting level, not level as is a level of a factor, right? In Figure 7
 (right), the nested bar charts are all sized proportionally with respect
 to each other even though they occur in different levels of the marital
 status factor.

 Figure 8 middle would be better with a bar chart, right? Why show both
 the middle and right plots; they're basically the same.

 Section 6.2

 Why limit yourself to a finite number of bins? Push the number of bins to
 infinity and you can display line graphs (density plots) with hbar,
 stacked line graphs (hbar+vspine), relative area (hspine+vspine),
 overlaid line charts (hbar+hbar), or (hspine+hbar) (don't know the name
 of this chart, basically overlaid line charts, but where the sum of
 values at each vertical slice is 1).

# Review 3

 I had some problems, though, in classifying this paper - it falls
 somewhere between theory/model (describes a theoretical framework that
 encompasses a large number of existing visualisations) and technique
 (code is provided to implement the framework). I'm not sure if it'd be
 better to simplify the classification by omitting mention of the code
 except as a note near the end - while it's interesting, it's not core to
 the paper.

 Doing this would leave more space to address some other minor concerns
 with the figures: in figures 2, 3, 5, 6 and 7 it's quite clear that the
 labels on the plot have spacing issues that could be resolved with larger
 figures. Additionally, figure 4 spills over the gutter into the text in
 the right column.

 In terms of text, I noticed some typos (first page, "This framework is
 based on the fundamental statistical idea that high any high dimensional
 distribution" has too many "high"s), one slight miss in phrasing ('weed
 down' sounds like a mixed metaphor between 'weed out' and 'cut down') and
 perhaps the last column of the conclusion could benefit from editing
 (repetition of 'generally' stands out).

# Minor changes

* There is a line break after "Abstract-". The abstract text should begin
on the same line.

* Figure 4 overruns its column.

* Page 2. "Uniformly distributed" suggest to me a uniform random distribution. Is there another way to say this?

* Figure 3. Left and Right are Top and Bottom. LaTeX refused to obey your
orders, as usual ;o)

* Please italicize n in the fourth line of Section 3. Check that other
parameters are properly italicized in the text.

* Page 1. "This framework is based on the fundamental statistical idea that
 high any high dimensional" Delete the first "high."

 "51 020 and 5 000" use comma delimiters.

*  Figure 12. The bottom row is not produced by transforming after rotation.
 It is produced by switching the assignment of x and y to r and theta.
 This is a pivoting operation.