### 2011-06-23 reviews/response.txt
More on R package.  Start on response

We are grateful for the helpful feedback provided by the reviewers. We have added a related work section, including many of the missing references we missed. We have also weakened our claims about the superiority of rectangles, instead providing a more pragmatic justification

Other changes:

* Rather than eliminating the section describing the R implementation altogether, we have expanded it slightly to give a taste of how it might be used.

------

# Review 1
 Finally, I wonder if the authors considered any other graphical
 composition method than nesting areas? While nesting areas allowed them
 to encompass a number of existing plot types, other composition methods
 might be more useful for extension to higher dimensions. For example,
 what about simply showing a separate plot for each component?

 -label all appropriate plots in the paper with the corresponding
 distribution decomposition (directly label, not just mention in the text
 or in the caption)
 -flesh out section 3.1. Assuming that infovis readers will not know what
 conditional and marginal distributions are, the authors should
 demonstrate how they are computed from the joint distribution. Providing
 the graphical analogs to the tables in Figure 4 would be extremely
 helpful. (I realize that a marginal and a conditional distribution from
 Figure 4 are represented in Figure 5, but the connection is never made
 explicitly in the text. Some hand holding would be very helpful here.)
 Hand hold better the analytic questions that each distribution helps
 answer. E.g. "What proportion of males are very happy? The bottom left
 entries tells us it's 31%."

 Minor comments----

 Why squarified treemaps? Heer work showing non-square treemaps are easier
 to read.

 Section 3.1

 Its not clear from the wording in the first paragraph that the marginal
 and conditional distributions contain less information than the joint
 distribution. If you're going to assume that your readers don't know what
 the marginal and conditional distributions are (plausible for the InfoVis
 crowd), then it would make sense to briefly describe how they are
 computed from the joint distribution.

 Fix the formatting of the tables in Figure 4.

 Clarify in the last paragraph that in addition to the 3 factorings,
 permutations of the variables leads to more variations.

 "It remains to find an [sic] graphical analog of this multiplication."
 seems unnecessarily dramatic. We already know you found one.

 Section 3.2

 Label Figure 5 with the distribution forms, f(happy), f(happy|sex),
 f(happy, sex), rather than just mention them in the text.

 Figure 7 shows, f(marital | age) f(age) and f(age | marital) f(marital),
 right? Say this in the text and label figures.

 "the relationship between proportion and area is only constant within a
 level" -- is there a better way to say this? I think you mean level as in
 nesting level, not level as is a level of a factor, right? In Figure 7
 (right), the nested bar charts are all sized proportionally with respect
 to each other even though they occur in different levels of the marital
 status factor.

 Figure 8 middle would be better with a bar chart, right? Why show both
 the middle and right plots; they're basically the same.

 Section 6.2

 Why limit yourself to a finite number of bins? Push the number of bins to
 infinity and you can display line graphs (density plots) with hbar,
 stacked line graphs (hbar+vspine), relative area (hspine+vspine),
 overlaid line charts (hbar+hbar), or (hspine+hbar) (don't know the name
 of this chart, basically overlaid line charts, but where the sum of
 values at each vertical slice is 1).

# Review 3

 I had some problems, though, in classifying this paper - it falls
 somewhere between theory/model (describes a theoretical framework that
 encompasses a large number of existing visualisations) and technique
 (code is provided to implement the framework). I'm not sure if it'd be
 better to simplify the classification by omitting mention of the code
 except as a note near the end - while it's interesting, it's not core to
 the paper.

 Doing this would leave more space to address some other minor concerns
 with the figures: in figures 2, 3, 5, 6 and 7 it's quite clear that the
 labels on the plot have spacing issues that could be resolved with larger
 figures. Additionally, figure 4 spills over the gutter into the text in
 the right column.

 In terms of text, I noticed some typos (first page, "This framework is
 based on the fundamental statistical idea that high any high dimensional
 distribution" has too many "high"s), one slight miss in phrasing ('weed
 down' sounds like a mixed metaphor between 'weed out' and 'cut down') and
 perhaps the last column of the conclusion could benefit from editing
 (repetition of 'generally' stands out).

# Minor changes

* There is a line break after "Abstract-". The abstract text should begin
on the same line.

* Figure 4 overruns its column.

* Page 2. "Uniformly distributed" suggest to me a uniform random distribution. Is there another way to say this?

* Figure 3. Left and Right are Top and Bottom. LaTeX refused to obey your
orders, as usual ;o)

* Please italicize n in the fourth line of Section 3. Check that other
parameters are properly italicized in the text.

* Page 1. "This framework is based on the fundamental statistical idea that
 high any high dimensional" Delete the first "high."

 "51 020 and 5 000" use comma delimiters.

*  Figure 12. The bottom row is not produced by transforming after rotation.
 It is produced by switching the assignment of x and y to r and theta.
 This is a pivoting operation.