### 2011-06-23 reviews/response.txt
More minor changes

We are grateful for the helpful feedback provided by the reviewers. We have added a related work section, including many of the missing references we missed. We have also weakened our claims about the superiority of rectangles, instead providing a more pragmatic justification

Other changes:

* Rather than eliminating the section describing the R implementation altogether, we have expanded it slightly to give a taste of how it might be used.

* Where appropriate, figures labelled with distributional decomposition.

* We wish we could make the figures larger and the labels easier to read, but the page limits are extremely tight this year.  We have done our best to ensure that all information is convey redundantly through colour or description in the captions.

------

# Review 1
Finally, I wonder if the authors considered any other graphical
composition method than nesting areas? While nesting areas allowed them
to encompass a number of existing plot types, other composition methods
might be more useful for extension to higher dimensions. For example,
what about simply showing a separate plot for each component?

-flesh out section 3.1. Assuming that infovis readers will not know what
conditional and marginal distributions are, the authors should
demonstrate how they are computed from the joint distribution. Providing
the graphical analogs to the tables in Figure 4 would be extremely
helpful. (I realize that a marginal and a conditional distribution from
Figure 4 are represented in Figure 5, but the connection is never made
explicitly in the text. Some hand holding would be very helpful here.)
Hand hold better the analytic questions that each distribution helps
answer. E.g. "What proportion of males are very happy? The bottom left
entries tells us it's 31%."

Minor comments----

Why squarified treemaps? Heer work showing non-square treemaps are easier
to read.

Section 3.1

Its not clear from the wording in the first paragraph that the marginal
and conditional distributions contain less information than the joint
distribution. If you're going to assume that your readers don't know what
the marginal and conditional distributions are (plausible for the InfoVis
crowd), then it would make sense to briefly describe how they are
computed from the joint distribution.

Fix the formatting of the tables in Figure 4.

Clarify in the last paragraph that in addition to the 3 factorings,
permutations of the variables leads to more variations.

"It remains to find an [sic] graphical analog of this multiplication."
seems unnecessarily dramatic. We already know you found one.

Section 3.2

Figure 7 shows, f(marital | age) f(age) and f(age | marital) f(marital),
right? Say this in the text and label figures.

"the relationship between proportion and area is only constant within a
level" -- is there a better way to say this? I think you mean level as in
nesting level, not level as is a level of a factor, right? In Figure 7
(right), the nested bar charts are all sized proportionally with respect
to each other even though they occur in different levels of the marital
status factor.

Figure 8 middle would be better with a bar chart, right? Why show both
the middle and right plots; they're basically the same.

Section 6.2

Why limit yourself to a finite number of bins? Push the number of bins to
infinity and you can display line graphs (density plots) with hbar,
stacked line graphs (hbar+vspine), relative area (hspine+vspine),
overlaid line charts (hbar+hbar), or (hspine+hbar) (don't know the name
of this chart, basically overlaid line charts, but where the sum of
values at each vertical slice is 1).

# Review 3

I had some problems, though, in classifying this paper - it falls
somewhere between theory/model (describes a theoretical framework that
encompasses a large number of existing visualisations) and technique
(code is provided to implement the framework). I'm not sure if it'd be
better to simplify the classification by omitting mention of the code
except as a note near the end - while it's interesting, it's not core to
the paper.

Doing this would leave more space to address some other minor concerns
with the figures: in figures 2, 3, 5, 6 and 7 it's quite clear that the
labels on the plot have spacing issues that could be resolved with larger
figures. Additionally, figure 4 spills over the gutter into the text in
the right column.

# Minor changes

* Figure 4 overruns its column.