### 2011-06-27 reviews/response.txt
Final tweaks

We are grateful for the helpful feedback provided by the reviewers. We have added a related work section, including all of the references we missed. We have also weakened our claims about the superiority of rectangles, instead providing a more pragmatic justification

Other changes:

* Rather than eliminating the section describing the R implementation altogether, we have expanded it slightly to give a taste of how it might be used.

* Figures are labelled with distributional decomposition.

* We have expanded the section that introduces joint, marginal and conditional densities to be more helpful for newcomers to the topic.

* We wish we could make the figures larger and the labels easier to read, but the page limits are extremely tight this year.  We have done our best to ensure that all information is convey redundantly through colour or description in the captions.

* We have endeavoured to make all of the other minor improvements to spelling and grammar that the reviewers suggested.  Thanks!